Staged Abstract Interpreters

### 2019-04-06 paper/70_eval.tex
Grammar eval

To evaluate the performance can be improved by our staging abstract interpreters
approach, we scale our toy interpreter to a large subset of the Scheme language
and evaluate the analyzing time against with the unstaged versions. The result
shows that the staged versions are averagely 6 times as fast of the unstaged
version, as a strong evidence of our \textit{abstraction without regret}
approach.

\paragraph{Implementation and Evaluation Environment}
We implement the abstract interpreters for control-flow analysis, i.e., a
value-flow analysis focus on lambda terms in functional languages. A front-end
desugars Scheme to a small core language that make the analyzer easy to
implement. As described in Section \ref{cfa}, we implement a monovariant
0-CFA-like analysis, \todo{a context-sensitive $1$-CFA-like analysis, both their
  versions with and without store-widening}. We also implement several
optimizations mentioned in Section \ref{staged_ds} (less than ~30 LOC in total).

We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores (192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on 0CFA without store-widening. The
\textit{unstaged} and \textit{staged} columns show the median time to finish the
analysis (in seconds). The last column $\frac{\text{unstaged}}{\text{staged}}$
shows the improvement from staging. \todo{0cfa, 1cfa, store-widen}
As we can see from the table, the staged version significantly outperforms the
unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
We observe the most improvement of 17x times faster than the unstaged version on
@kcfa-worst-64@. On average, the staging produces code that run 7x times faster
than the unstaged analyzer.

We also observe that much overhead of unstaged analyzers are from the monadic
layers. Thus our evaluation result provides an abstraction without regret
approach to construct optimizing abstract interpreters -- user may write a
high-level modular and composable abstract interpreter, and then derive a staged
version in a principled way that runs significantly faster.

>>>>

To evaluate how the performance can be improved by our staged abstract interpreters
approach, we scale our toy interpreter to a large subset of the Scheme language
and evaluate the analyzing time against with the unstaged versions. The result
shows that the staged versions are averagely 6 times as fast of the unstaged
version, as strong evidence of our \textit{abstraction without regret}
approach.

\paragraph{Implementation and Evaluation Environment}
We implement the abstract interpreters for control-flow analysis, i.e., a
value-flow analysis focus on lambda terms in functional languages. A front-end
desugars Scheme to a small core language that makes the analyzer easy to
implement. As described in Section \ref{cfa}, we implement a monovariant
0-CFA-like analysis, \todo{a context-sensitive $1$-CFA-like analysis, both their
  versions with and without store-widening}. We also implement several
optimizations mentioned in Section \ref{staged_ds} (less than ~30 LOC in total).

We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
(kernel 4.4.0) machine. All of our evaluations were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores (192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on 0CFA without store-widening. The
\textit{unstaged} and \textit{staged} columns show the median time to finish the
analysis (in seconds). The last column $\frac{\text{unstaged}}{\text{staged}}$
shows the improvement from staging. \todo{0cfa, 1cfa, store-widen}
As we can see from the table, the staged version significantly outperforms the
unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
We observe the most improvement of 17x times faster than the unstaged version on
@kcfa-worst-64@. On average, the staging produces code that runs 7x times faster
than the unstaged analyzer.

We also observe that much overhead of unstaged analyzers are from the monadic
layers. Thus our evaluation result provides an abstraction without regret
approach to construct optimizing abstract interpreters -- the user may write a
high-level modular and composable abstract interpreter, and then derive a staged
version in a principled way that runs significantly faster.
